---
title: "R Introduction and SDT"
author: "Dr. Zhuanghua Shi"
date: "28 Nov. 2016"
output: beamer_presentation
header-includes:
  - \usetheme{metropolis}
  - \def\begincols{\begin{columns}\begin{column}}
  - \def\endcols{\end{column}\end{columns}}
  - \def\colsep{\end{column}\begin{column}}
---
  
```{r setup, include=FALSE}
library(knitr)
library(pander)
library(dplyr)
library(ggplot2)
library(xtable)
library(fitdistrplus)
library(ez)
library(statmod) # invgauss distribution
options(xtable.comment=FALSE, digits = 2)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, fig.align='center')
```

## Last week homework (1)

```{r, echo = TRUE}
dat = read.csv('search_rt.csv')
# summary(dat) # use this to check the data structure
# find out outliers
msub <- dat %>% filter(error == FALSE) %>% 
  group_by(sub) %>% 
  summarise(sd_rt = sd(rt), mrt = mean(rt)) %>%
  mutate(low_bound =  mrt - 2.5*sd_rt, 
         up_bound = mrt+2.5*sd_rt)
```

## Last week homework (2)

```{r, echo = TRUE, results = 'asis'}
# average reaction times for individual subjects
msub_rt <- left_join(dat, msub, by = c('sub')) %>% 
  mutate(outlier = rt>up_bound | rt<low_bound) %>%
  filter(outlier == FALSE & error == FALSE) %>% 
  group_by(sub, BlkType, target) %>% 
  summarise(mrt = mean(rt), sd = sd(rt))
#summary(msub_rt)
xtable(msub_rt[1:5,])
```

## Last week homework (3)

```{r, echo = TRUE, fig.width= 4, fig.height= 3, out.width= '50%'}
# plot graphs
msub_rt %>% group_by (BlkType, target) %>% 
summarise(mmrt = mean(mrt), se = sd(mrt)/sqrt(nlevels(sub)-1)) %>%
ggplot(aes(x = BlkType, y = mmrt, fill = target, color = target)) + 
geom_bar(stat = 'identity', position = 'dodge') + 
geom_errorbar(aes(ymax = mmrt + se, ymin = mmrt - se), 
position = 'dodge') + coord_cartesian(ylim = c(0.4, 0.6))

```

## Last week homework (4)

```{r, echo = TRUE, results='asis'}

anova_result <- ezANOVA(data.frame(msub_rt),
                        dv = mrt, wid = sub, 
                        within = .(BlkType, target))
xtable(anova_result$ANOVA)
```


## Last week homework (5)
```{r, echo = TRUE, fig.width=4, fig.height= 3, out.width='50%'}
# plot using ez package
ezPlot(data.frame(msub_rt), dv = mrt, wid = sub, 
       within = .(BlkType, target), x = BlkType, split = target)

```

## Signal Detection Theory 

* Signal Detection table

Response    Signal    Noise
--------    -------   -----------------------
Yes         Hit(H)    False Alarm (FA)
No          Miss(M)   Correct rejection (CR)

* Eaqual variance 

$$ d' = z(H) - Z(FA)$$

$$ C = - \frac{(z(H)+ z(FA))}{2}$$

## Example A: Multiple-choice Test

`sub_grades` was from a multiple-choise test made by students. In total there were 10 correct items and 10 incorrect items. The table shows individuals' performance. Your task is to calculate the d' and C.  

```{r, echo = TRUE}
grades <- read.csv2('sub_grades.csv')
# examin the data and refine
grades <- grades[1:13,1:5]

```

## Example A: Calculate d' and C


```{r, echo = TRUE}
# Calculate d', and C
grades <- grades %>% mutate(H = Hit/Correct.Items, 
                  FA = 1- Correct.Rejection/Incorrect.Items) %>%
  mutate(zH = qnorm(pmin(H,0.999)), zF = qnorm(pmax(0.001, FA))) %>%
  mutate(d = zH-zF, C = -(zH+zF)/2)

```

## Example A: Visualize results

```{r, echo = TRUE, fig.width=4,fig.height=3, out.width= '60%'}

ggplot(grades, aes(x = d, y = C, label = StudNo)) + 
  geom_point() + geom_text(hjust = 0, nudge_x = 0.05)

```

## Example A: Visualize results (2)

```{r, echo = TRUE, fig.width=4,fig.height=3, out.width= '60%'}

ggplot(grades, aes(x = FA, y = H, label = StudNo)) + 
  geom_point() + geom_text(hjust = 0, nudge_x = 0.05) + 
  coord_cartesian(xlim = c(0,1), ylim = c(0,1))

```

## Example B: Confidence rating judgment

```{r, echo = TRUE, results = 'asis'}

rating <- read.csv2('confidence_rating.csv')

xtable(rating)
```

## Example B: Your task

* Calculate d' and C for different confidence levels
* Plot a ROC curve

